### 2016-06-29 cover_letter.tex
Add cover letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thin Formal Letter
% LaTeX Template
% Version 1.11 (8/12/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (http://en.wikibooks.org/wiki/LaTeX/Letters)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{letter}

% Adjust margins for aesthetics
\addtolength{\voffset}{-0.5in}
\addtolength{\hoffset}{-0.3in}
\addtolength{\textheight}{2cm}

%\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," to the left of the page

%----------------------------------------------------------------------------------------
%YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\signature{G. Lemaitre, F. Nogueira, and C. Aridas} % Your name for the signature at the bottom

\address{12 rue de la Fonderie \\ 71200 Le Creusot \\ +33(0)761104782} % Your address and phone number

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Editor of JMLR} % Name/title of the addressee

%----------------------------------------------------------------------------------------
%LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{\textbf{Dear Sir or Madam,}}

This submission is intended for the machine learning open source software section.
Our software is a Python library to tackle the curse of imbalanced datasets in machine learning.
The recent popularity of this project strengthen us in our decision of disseminating the proposed package. 
This toolbox is released under an open source MIT license and is hosted in GitHub:

\texttt{https://github.com/fmfn/UnbalancedDataset}

The attached version to be reviewed is the \texttt{0.1} release and is also attached in a \texttt{tar.gz} archive.
A five pages description of the software is also provided emphasizing on the benefit of this software for machine learning application as well as the project vision.

Thank you for your time and consideration.

We look forward to your reply.

\vspace{2\parskip} % Extra whitespace for aesthetics
\closing{Sincerely,}
\vspace{2\parskip} % Extra whitespace for aesthetics

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}